Microbial interactions as drivers of a nitrification process in a chemostat

### 2020-05-26 tex/Article_Interactions.tex
modofications to sup mat

\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{algorithm2e}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage[version=4]{mhchem}
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\I}{\mathcal{I}}
\DeclareMathOperator{\diag}{diag}
\newtheorem{defn}{Definition}
\newtheorem{hypo}{H}
\newtheorem{lemma}{Lemma}
\newtheorem{theo}{Theorem}
\usepackage[figuresright]{rotating}

@@

\begin{align} 
\label{system}
\begin{array}{cl}
\dot{x_i} =& \left(\mu_i(s,x) -D \right)x_i \quad \forall i \in [n] \\
\dot{s_1} =& \displaystyle (s_{in}-s_1)D+\sum\limits_{i\in G_1} y_{s_1/x_i}\mu_i(s,x) x_i  \\
\dot{s_2} = & \displaystyle -s_2D+\sum\limits_{i\in [n]}y_{s_2/x_i}\mu_i(s,x)x_i  \\
\dot{s_3} =&  \displaystyle -s_3D+\sum\limits_{i\in G_2}y_{s_3/x_i}\mu_i(s,x) x_i 
\end{array}
\end{align}

@@

\begin{lemma}
	\label{l1}
	For initial conditions $(x_1(0),\dots,x_n(0),s_1(0),s_2(0),s_3(0)))\in \Omega$, there exists positive scalars $M_1$, $M_2$, and $M_3$ such that solutions to \eqref{system} satisfy the following inequalities:
	\begin{align}
	&\sum \limits_{i \in G_1} \frac{1}{y_i}x_i + s_1 \leq M_1 \\
	&\sum \limits_{i \in G_2} \frac{1}{y_i}x_i + s_1 +s_2 \leq M_2 \\
	&s_1 + s_2 + s_2 \leq M_3
	\end{align}
\end{lemma}

@@

\begin{theo}
	\label{theoWellPosedness}
	For initial conditions $(x_1(0),\dots,x_n(0),s_1(0),s_2(0),s_3(0)))\in \Omega$, there exists a constant $M>0$ such that for every matrix $A$ satisfying $\Vert A \Vert_{\infty} \leq M $, the solutions of system \eqref{system} with growth rates given by \eqref{gLV growth1} and \eqref{gLV growth2}, remain in $\Omega$ and are bounded.
\end{theo}

@@

In this section analytical expressions for equilibrium points are derived. However, no analytic expression concerning the stability of such points is presented. To answer the question of stability a numeric scheme is used by evaluating the Jacobian at the equilibrium point. At the end of the section an algorithm is provided for exploring all the possible equilibria.

@@

\begin{defn}
	Given an equilibrium point $(x^{eq},s^{eq})$ of system \eqref{system}, then the set of non-active coordinates $\mathcal{J}\subset \{1,\dots,n\}$ is defined as:
	$\mathcal{J} = \{j_1, \dots, j_m\ : \, x^{eq}_{j_i} = 0 , \;i \in [m] \} $. $n_1^{act}$ and $n_2^{act}$ denote the number of positive entries of $x^{eq}$ of functional groups $G_1$ and $G_2$, respectively. $n^{act} = n-m$ denotes the total number of positive entries of $x^{eq}$. The active point $x^{act}\in \R^{n^{act}}$ is defined by the positive entries of $x^{eq}$. Analogously, the functions $f^{act}(s)$ and $\mu^{act}(x,s)$ are defined by the positive entries of $x^{eq}$. The active interactions $A^{act}$ is defined as the matrix $A$ without the $\mathcal{J}$ rows and columns. The active stoichiometry matrix $Y^{act}$ is the matrix $Y$ without the $\mathcal{J}$ columns. In summary, let $M$ be the matrix defined by taking out the $\mathcal{J}$ columns of the identity matrix of size $n$, then
	 \begin{align}
	 x^{eq} = Mx^{act} \\	
	 f^{act}(s) = M^{\top}f(s) \\
	 \mu^{act}(x,s) = M^{\top}\mu(x,s)\\
	 Y^{act} = YM \\
	 A^{act} = M^{\top}AM
	 \end{align}.   
\end{defn}

 Finally, for deriving the equilibrium points, it is desirable an invertible $A^{act}$ matrix. Therefore in what follows of the work it is assumed that matrix $A$ and some of its sub matrices have an inverse, this is stated properly in Hypothesis \ref{inverse hypothesis}.

@@

@@

note that $\forall j \in \mathcal{J}, \; x^{eq}_j = 0 $ therefore the coefficients $a_{ij}$ play no role in equation \eqref{Extinction} and one can rewrite equation \eqref{Extinction} as $\diag(f^{act}(s))(1_{n^{act}\times 1} + A^{act}x^{act}) - D_{n^{act} \times 1}=0 $. From hypothesis \ref{inverse hypothesis} matrix $A^{act}$ has an inverse. This gives the following lemma.

\begin{lemma} 
	The active point of the solution of \eqref{Extinction} solves the system of equations \eqref{EqSome}. 
	\begin{align}
	\label{EqSome}x^{act} = (A^{act})^{-1}(\diag(f^{act}(s))^{-1}D_{n^{act}\times 1} - 1_{n^{act}\times 1})
	\end{align} 
\end{lemma}

By replacing $x^{eq} = Mx^{act}$ in equation \eqref{Eq2}:

\begin{align}
 (s_{in}-s_1)D + Y_{1\bullet} \diag(\mu(x,s))Mx^{act}  = 0 \\
(s_{in}-s_1)D + Y^{act}_{1\bullet} \diag(\mu^{act}(x,s))x^{act}  = 0 \\ 
(s_{in}-s_1)D + Y^{act}_{1\bullet} \diag(D_{n^{act}\times 1})x^{act}  = 0  \\
(s_{in}-s_1) + Y^{act}_{1\bullet} x^{act}  = 0  \\
\label{s1Equilibria} (s_{in}-s_1) + Y^{act}_{1\bullet}(A^{act})^{-1}(\diag(f^{act}(s))^{-1}D_{n^{act}\times 1} - 1_{n^{act}\times 1}) = 0
\end{align}

A similar expression is obtained when replacing \eqref{EqSome} in equation \eqref{Eq3}

\begin{align}
\label{s2Equilibria}-s_2+ Y^{act}_{2\bullet}(A^{act})^{-1}(\diag(f^{act}(s))^{-1}D_{n^{act}\times 1} - 1_{n^{act}\times 1}) = 0
\end{align}

From equation \eqref{s1Equilibria} $s_2$ can be written as a function of $s_1$:
\begin{align}
s_2 = \frac{s_1}{c_1s_1^2+c_2s_1+c_3} \label{S2(S1)}
\end{align}

Then by replacing  \eqref{S2(S1)} in equation \eqref{s2Equilibria}, one gets a fourth degree polynomial for $s_1$.

@@

Finally, in equilibrium, $s_3$ can be defined in terms of $s_1$, $s_2$ and $s_{in}$. This is done by adding equations \eqref{Eq2}, \eqref{Eq3},  and \eqref{Eq4} which gives :

If the system of equations \eqref{EqSome}, \eqref{S2(S1)}, \eqref{Poly4}, and \eqref{s1s2s3} provides only solutions with some negative entries then the set $\mathcal{J}$ can not define an equilibrium point.

@@

Equations \eqref{EqSome} and \eqref{s1Equilibria} still hold, but in this case note that $f^{act}(s)$ does not contain any term involving $s_2$, and therefore one obtains a quadratic equation for $s_1$:

@@

Since $x_i = 0, \, \forall i \in G_2$ then from equation \eqref{Eq4}.
\begin{align}
\dot{s_3} = 0 = -s_3D \\
\Rightarrow s_3 = 0
\end{align}

Then by \eqref{s1s2s3} one obtains:

\begin{align}
s_{in} = s_1 + s_2
\end{align}

@@

From equation \eqref{Eq2}, one gets 
\begin{align*} s_{in} = s_1 \end{align*}
then \eqref{s1s2s3} implies \begin{align*}s_2 &= s_3 = 0 \\ \Rightarrow \dot{x_i} &= -Dx_i \quad \forall i = n_1+1, \dots, n\end{align*}   
which has an equilibria at $x_i = 0$.

@@

\begin{table}[ht]
	\centering
	\begin{tabular}{|l|l|l|l|}
		\hline
		Kinetic Parameters & $\mu_i\,[1/day]$ & $K_i\,[g/L]$ & $\frac{1}{y_i} \, [gr/gr]$ \\ \hline
		$x_1 \in G_1$ & 0.77  & 0.7 & 3.98  \\ \hline
		$x_2\in G_2$ & 1.07 & 0.3 &  16.12 \\ \hline
	\end{tabular}	
	\caption{A set of kinetic parameters of model \eqref{system}.}
	\label{kinetic_parameters_case_study_1}
\end{table}

@@

 \begin{table}[ht]
 	\centering
 	\begin{tabular}{|l|l|l|l|}
 		\hline
 		Case study 2 kinetic parameters & $\mu_i\,[1/day]$ & $K_i\,[mg/L]$ & $\frac{1}{y_i} \, [gr/gr]$ \\ \hline
 		$x_1 \in G_1$ & 0.828  & 0.147 & 3.85  \\ \hline
 		$x_2\in G_1$ &0.828   & 0.147 &  3.85\\ \hline
 		$x_3\in G_2$ & 0.18 & 0.026 &  100 \\ \hline
 		$x_4\in G_2$ & 0.18 & 0.026 &  100 \\ \hline
 	\end{tabular}
 	\caption{Kinetic parameters of model \eqref{system} from Dumont \textit{et al.} \cite{Dumont2016}}
 	\label{kinetic_parameters}
 \end{table}

@@

\begin{align} 
\label{Controlsystem}
\begin{array}{cl}
\dot{x_i} =& \left(f_i(s)u_i(t) -D \right)x_i \quad \forall i \in G_1\\
\dot{x_i} =& \left(f_i(s)u_i(t) -D \right)x_i \quad \forall i \in G_2\\
\dot{s_1} =& \displaystyle (s_{in}-s_1)D-\sum\limits_{i \in G_1}\frac{1}{y_i}f_i(s)u_i(t) x_i  \\
\dot{s_2} = & \displaystyle -s_2D+\sum\limits_{i \in G_1}\frac{1}{y_i}f_i(s)u_i(t)	-\sum\limits_{i \in G_2}\frac{1}{y_i}f_i(s)u_i(t) x_i  \\
\dot{s_3} =&  \displaystyle -s_3D+\sum\limits_{i \in G_2}\frac{1}{y_i}f_i(s)u_i(t) x_i \\
y  =& x
\end{array}
\end{align}

 Consider the weighted norms defined by positive definite matrices $Q$ and $R$, represented by $\Vert \cdot \Vert_Q$ and $\Vert \cdot \Vert_R$, respectively, and a positive real $\bar{u}$. The optimal tracking problem is defined as:

@@

\item Second, to force a regularized control. Otherwise note that $u$ is linear in \eqref{Controlsystem}, therefore if the integral cost does not have a non-linear expression of $u$ the optimal control will be of a bang-bang type with possibly singular arcs \cite{harmand2019optimal}. Since the objective is to find a differentiable expression of $I(x)$ the addition of the regularization term is deemed necessary.
\end{itemize} 
The problem of approximating the solution of the system to a desired reference ($z$ in this case) is called the optimal tracking problem. For solving such a problem the approach developed by Cimen \textit{et al.} \cite{Cimen2004, Cimen2008} was adapted to our problem. The method proposed involves the resolution of Approximating Sequences of Ricatti Equations (ASRE).

The method consists on iteratively calculating trajectories of System \eqref{Controlsystem} with a certain control and feeding a non-autonomous Ricatti differential equation with the resulting trajectory. Then, a control law that uses the solution of the Ricatti equation is proposed and a new trajectory is calculated. The iteration is stopped when a convergence in the output or  the control is observed.

The tracking problem does not consider a constrained control. Nevertheless, the methods of Cimen \textit{et al.}\cite{Cimen2004} were directly used with an explicit constraint in the synthesis of the control. Even though this is probably suboptimal when the control reaches its bounds, one at least is certain about its optimality when the control never reaches its constraints.

@@

The change of variable $u_i(t) = v_i(t) + 1$ is used for technical reasons explained in the supplementary material section. This in turn implies $v_i(t) \in [-1,\bar{u}-1]$.

The feedback control $v(t)$, will be of the form $v(t) = -R^{-1}\tilde{B}^\top(t)\left(\tilde{P}(t)x(t)-s_f(t)\right)$. Where matrix $P(t)$ and vector $s_f(t)$ solve differential equations. In the appendix it is proved that because of the structure of the system one only needs to calculate $2n$ differential equations for the synthesis of the control, instead of $(n+3)^2 + (n+3)$ that would imply the direct application of the method, which renders the method- at least theoretically- scalable for a growing number of OTU.

@@

The tracking problem was applied to data coming from a nitrification process with experimental conditions described in \cite{dumont2008observers}. For exploring the hypothesis of interactions as drivers of bioreactors performance environmental conditions should be kept as constant as possible. Therefore only data from day $183$ onwards was used because a change in the operating temperature happened at that point, which is known to have an effect on kinetics. For choosing which species belong in which functional group, the procedure described in \cite{Ugalde-Salas2019} was used: from day $183$ to day $315$: 31 OTU were identified in the $G_1$ group (AOB) and 5 in the $G_2$ functional group (NOB).

@@

The knowledge of functions $f_i(s)$ was based on a study of nitrification's kinetic parameters\cite{Wiesmann1994}. Particularly given the system's ammonium and nitrate concentration a Monod function (eq \eqref{monod}) was used for $G_1$ and $G_2$ with parameters given in table \ref{kinetic_parameters_application}. The yields were fitted to match the nitrogen mass balances. The $Q$ and $R$ matrices were the same as in the proof of concept section, that is $\begin{bmatrix}
\lambda_1 I_{n_1} &0  \\ 0& \lambda_2 I_{n_2}
\end{bmatrix}$ and $I_n$, respectively, with $\lambda_1 = 10^{-4}$ and $\lambda_2 = 10^{-5}$.

@@

\begin{table}[ht]
	\centering
	\begin{tabular}{|l|l|l|l|}
		\hline
		Kinetic Parameters & $\mu_i\,[1/day]$ & $K_i\,[g/L]$ & $\frac{1}{y_i} \, [gr/gr]$ \\ \hline
		$x_1 \in G_1$ & 1.97  & $7\cdot 10^{-1}$ & 4.49  \\ \hline
		$x_2\in G_2$ & 1.87 & $5.4\cdot 10^{-1}$ &  45.51 \\ \hline
	\end{tabular}	
	\caption{A set of kinetic parameters of model \eqref{system}.}
	\label{kinetic_parameters_application}
\end{table}
For the reader to gain understanding of the situation, a simulation of the system using the experiments operating parameters ($D$ and $s_{in}$) is presented without control (i.e. $u(t)= 1$) in  figure \ref{u=1 simulation.} nitrite ($s_3$) accumulates all along the trajectory, but when compared to data it is clear that $s_3$ stops accumulating after a while.

@@

Last decades advances in genetic sequencing and microbial ecology have opened a gap for modellers in biochemical processes to integrate this valuable information. Considering the success of mass balance models to predict and pilot bioreactors, one should build upon these existing models. This article proposed a naïve approach to start filling the gap of microbial ecology and biochemical process modelling: The approach can be simply described as correcting the growth rate of each individual species from genetic sequencing measurements, and observing what would that imply according to the (pseudo)stoichiometric equations involved. If one already posses a model for describing a particular bioprocess the use of the techniques here presented can be applied in a straightforward manner if one has already identified the microbial species involved with a particular functionality of the system.

@@

\textbf{Proof of lemma \ref{l1}}
\begin{proof}
	Define 
	\begin{align}
	z_1 &: = \sum \limits_{i \in G_1} \frac{1}{y_i}x_i + s_1  \\
	z_2 &: = \sum \limits_{i \in G_2} \frac{1}{y_i}x_i + s_1 + s_2 \\
	z_3 &: = s_1 + s_2 + s_3
	\end{align}

@@

\textbf{Proof of theorem \ref{theoWellPosedness}}

@@

	If $s_2 = 0$ then note first that by adding both inequalities from Lemma \eqref{l1} and since, $s_1,x_i \geq 0$ one has:
	\begin{align*}
	\sum \limits_{i \in G_1 \cup G_2}  \frac{1}{y_i} x_i + 2s_1 + s_2 \leq M_1 + M_2 \\
	\Rightarrow \sum \limits_{i = 1}^{n}  \frac{1}{y_i} x_i + 2s_1 \leq M_1 + M_2 \\
	\Rightarrow \sum \limits_{i = 1}^{n}  \frac{1}{y_i} x_i \leq M_1 + M_2\\ 
	\Rightarrow \bar{k} \sum  \limits_{i = 1}^{n}   x_i  \leq M_1 + M_2  \\
	\Leftrightarrow \bar{k} \sum  \limits_{i = 1}^{n}   \vert x_i \vert  \leq M_1 + M_2  \\
	\Leftrightarrow \bar{k} \Vert x \Vert_1  \leq M_1 + M_2  \\
	\Rightarrow   \bar{k} \Vert x \Vert_{\infty} \leq M_1 + M_2
	\end{align*}

	Where $\bar{k} = \min \left\{ \frac{1}{y_i}  : \; i\in [n] \right \}$.

	Define $M := \dfrac{\bar{k}}{M_1 + M_2}$, and let $A$ be a matrix such that $\Vert A \Vert_{\infty} \leq M$. Then since $\Vert Ax \Vert_{\infty} \leq 1 \Rightarrow \forall i=1,\dots,n \; (1 + A_{i\bullet}x) \geq 0 $, and $\Vert Ax \Vert_{\infty} \leq \Vert A \Vert_{\infty} \Vert x\Vert_{\infty} \leq M \frac{1}{M}$.
	One has proven that:
	\begin{align}
	\label{Bound1} \forall i=1,\dots,n \;  \;  (1 + A_{i\bullet}x) \geq 0 
	\end{align}

@@

	For the boundedness it suffices to notice from Lemma \ref{l1} that the sum of positive elements is bounded, therefore each element is bounded.
\end{proof}

\subsection{Properties of diag operator}

Notice that for two vectors $v_1$ and $v_2$ the following holds:
\begin{align}
\diag(v_1)\diag(v_2) = \diag(v_2)\diag(v_1) 
\end{align}

Suppose $v \in \R^n$ such that $v_i \neq 0 \forall i \in [n]$. Define $v^{-1}:= \left( \dfrac{1}{v_1},\dots,\dfrac{1}{v_n} \right)$ then the inverse of $\diag(v)$ exists and is $\diag(v^{-1})$:

\begin{align}
\diag(v)^{-1} = \diag(v^{-1}) 
\end{align}

>>>>

\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{thmtools}
\usepackage{thm-restate}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{cleveref}
\usepackage{algorithm2e}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{xcolor}
\usepackage[version=4]{mhchem}
\newcommand{\N}{\mathbb{N}}
\newcommand{\R}{\mathbb{R}}
\newcommand{\I}{\mathcal{I}}
\DeclareMathOperator{\diag}{diag}
\declaretheorem[name=Theorem]{thm}
\declaretheorem[name=Lemma]{lemma}
\newtheorem{defn}{Definition}
\newtheorem{hypo}{H}
\usepackage[figuresright]{rotating}

@@

\begin{align} 
\label{system}
\begin{array}{cl}
\dot{x_i} =& \left(\mu_i(s,x) -D \right)x_i \quad \forall i \in [n] \\
\dot{s_1} =& \displaystyle (s_{in}-s_1)D - \sum\limits_{i\in G_1} \frac{1}{y_i} \mu_i(s,x) x_i  \\
\dot{s_2} = & \displaystyle -s_2D+\sum\limits_{i\in G_1} \frac{1}{y_i} \mu_i(s,x) x_i -\sum\limits_{i\in G_2} \frac{1}{y_i} \mu_i(s,x) x_i \\
\dot{s_3} =&  \displaystyle -s_3D+\sum\limits_{i\in G_2} \frac{1}{y_i} \mu_i(s,x) x_i  
\end{array}
\end{align}

Which can be written in a more compact notation using the stoichiometric matrix $Y$ and the $\diag$ operator.
\begin{align}
\label{Eq1X}	\dot{x} = & \diag(\mu(x,s) - D_{n\times 1})x \\
\label{EqS}	\dot{s} = & \left(\begin{bmatrix} s_{in}&0&0 \end{bmatrix}^\top-s\right)D + Y\diag(\mu(x,s))x
\end{align}

@@

\begin{restatable}{lemma}{systembounds}
	\label{l1}
	For initial conditions $(x_1(0),\dots,x_n(0),s_1(0),s_2(0),s_3(0)))\in \Omega$, there exists positive scalars $M_1$, $M_2$, and $M_3$ such that solutions to \eqref{system} satisfy the following inequalities:
	\begin{align}
	&\sum \limits_{i \in G_1} \frac{1}{y_i}x_i + s_1 \leq M_1 \\
	&\sum \limits_{i \in G_2} \frac{1}{y_i}x_i + s_1 +s_2 \leq M_2 \\
	&s_1 + s_2 + s_2 \leq M_3
	\end{align}
\end{restatable}

@@

\begin{restatable}{thm}{wellposedness}
	\label{theoWellPosedness}
	For initial conditions $(x_1(0),\dots,x_n(0),s_1(0),s_2(0),s_3(0)))\in \Omega$, there exists a constant $M>0$ such that for every matrix $A$ satisfying $\Vert A \Vert_{\infty} \leq M $, the solutions of system \eqref{system} with growth rates given by \eqref{gLV growth1} and \eqref{gLV growth2}, remain in $\Omega$ and are bounded.
\end{restatable}

@@

In this section analytical expressions for equilibrium points are shown. However, no analytic expression concerning the stability of such points is presented. To answer the question of stability a numeric scheme is used by evaluating the Jacobian at the equilibrium point. At the end of the section an algorithm is provided for exploring all the possible equilibria. All the computations for deriving the equations of this section can be found in the supplementary material.

@@

\begin{defn}
	Given an equilibrium point $(x^{eq},s^{eq})$ of system \eqref{system}, then the set of non-active coordinates $\mathcal{J}\subset \{1,\dots,n\}$ is defined as:
	$\mathcal{J} = \{j_1, \dots, j_m\ : \, x^{eq}_{j_i} = 0 , \;i \in [m] \} $. $n_1^{act}$ and $n_2^{act}$ denote the number of positive entries of $x^{eq}$ of functional groups $G_1$ and $G_2$, respectively. $n^{act} = n-m$ denotes the total number of positive entries of $x^{eq}$. The active point $x^{act}\in \R^{n^{act}}$ is defined by the positive entries of $x^{eq}$. Analogously, the functions $f^{act}(s)$ and $\mu^{act}(x,s)$ are defined by the positive entries of $x^{eq}$. The active interactions $A^{act}$ is defined as the matrix $A$ without the $\mathcal{J}$ rows and columns. The active stoichiometry matrix $Y^{act}$ is the matrix $Y$ without the $\mathcal{J}$ columns. 
\end{defn}

 In order to derive the equilibrium points, it is desirable an invertible $A^{act}$ matrix. Therefore in what follows of the work it is assumed that matrix $A$ and some of its sub matrices have an inverse, this is stated properly in Hypothesis \ref{inverse hypothesis}.

@@

Assuming Hypothesis \ref{inverse hypothesis} a formula for the active points is derived from equation \eqref{Eq1}:

\begin{align}
\label{EqSome}x^{act} = (A^{act})^{-1}(\diag(f^{act}(s))^{-1}D_{n^{act}\times 1} - 1_{n^{act}\times 1})
\end{align}

Note as well that at the equilibrium, $s_3$ can be defined in terms of $s_1$, $s_2$ and $s_{in}$. This is done by adding equations \eqref{Eq2}, \eqref{Eq3},  and \eqref{Eq4} which gives :

@@

By replacing equation \eqref{EqSome} in equation \eqref{Eq2} $s_2$ can be written as a function of $s_1$:
\begin{align}
s_2 = \frac{s_1}{c_1s_1^2+c_2s_1+c_3} \label{S2(S1)}
\end{align}

Then by replacing  \eqref{S2(S1)} in equation \eqref{Eq3}, one gets a fourth degree polynomial for $s_1$.

@@

The equilibrium point can be calculated from the solutions of the system of equations \eqref{EqSome}, \eqref{s1s2s3}, \eqref{S2(S1)}, and \eqref{Poly4} with non negative coordinates. If the system only provides solutions with at least one negative entry then the set $\mathcal{J}$ can not define an equilibrium point.

@@

Under this case note that $f^{act}(s)$ depends only on $s_1$. Therefore when equation \eqref{EqSome} is replaced in \eqref{Eq2}, one obtains a quadratic equation for $s_1$:

@@

Since $x_i = 0 \, \forall i \in G_2$ then from equation \eqref{Eq4}.
\begin{align}
\dot{s_3} = 0 = -s_3D \\
\Rightarrow s_3 = 0 \label{no_s3}
\end{align}

In this case the equilibrium point can be calculated from the solutions of the system of equations \eqref{EqSome}, \eqref{s1s2s3}, \eqref{poly2_s1}, and \eqref{no_s3} with non-negative coordinates. If the system only provides solutions with at least one negative entry then the set $\mathcal{J}$ can not define an equilibrium point.

@@

From equation \eqref{Eq2}, one gets 
\begin{align*} s_{in} = s_1 \end{align*}
then \eqref{s1s2s3} implies 
\begin{align*}s_2 &= s_3 = 0 \end{align*}

The equilibrium is then given by $\begin{pmatrix} 0_{1\times n}& s_{in}&0 &0 \end{pmatrix}^\top$.

@@

\begin{table}[ht]
	\centering
	\begin{tabular}{|l|l|l|l|}
		\hline
		Kinetic Parameters & $\bar{\mu}_i\,[1/day]$ & $K_i\,[g/L]$ & $\frac{1}{y_i} \, [gr/gr]$ \\ \hline
		$x_1 \in G_1$ & 0.77  & 0.7 & 3.98  \\ \hline
		$x_2\in G_2$ & 1.07 & 0.3 &  16.12 \\ \hline
	\end{tabular}	
	\caption{A set of kinetic parameters of model \eqref{system}.}
	\label{kinetic_parameters_case_study_1}
\end{table}

@@

 \begin{table}[ht]
 	\centering
 	\begin{tabular}{|l|l|l|l|}
 		\hline \rule{0pt}{3.5ex}
 		Case study 2 kinetic parameters & $\mu_i\,[1/day]$ & $K_i\,[mg/L]$ & $\frac{1}{y_i} \, [gr/gr]$ \\ \hline 
 		$x_1 \in G_1$ & 0.828  & 0.147 & 3.85  \\ \hline
 		$x_2\in G_1$ &0.828   & 0.147 &  3.85\\ \hline
 		$x_3\in G_2$ & 0.18 & 0.026 &  100 \\ \hline
 		$x_4\in G_2$ & 0.18 & 0.026 &  100 \\ \hline
 	\end{tabular}
 	\caption{Kinetic parameters of model \eqref{system} from Dumont \textit{et al.} \cite{Dumont2016}}
 	\label{kinetic_parameters}
 \end{table}

@@

\begin{align} 
\label{Controlsystem}
\begin{array}{cl}
\dot{x_i} =& \left(f_i(s)u_i(t) -D \right)x_i \quad \forall i \in G_1\\
\dot{x_i} =& \left(f_i(s)u_i(t) -D \right)x_i \quad \forall i \in G_2\\
\dot{s_1} =& \displaystyle (s_{in}-s_1)D + \sum\limits_{i \in G_1}y_{s_1/x_i}f_i(s)u_i(t) x_i  \\
\dot{s_2} = & \displaystyle -s_2D+\sum\limits_{i \in G_1 \cup G_2}y_{s_2/x_i}f_i(s)u_i(t)x_i  \\
\dot{s_3} =&  \displaystyle -s_3D+\sum\limits_{i \in G_2}y_{s_3/x_i}f_i(s)u_i(t) x_i \\
y  =& x
\end{array}
\end{align}

 Consider the weighted norms defined by positive definite matrices $Q$ and $R$, represented by $\Vert \cdot \Vert_Q$ and $\Vert \cdot \Vert_R$, respectively, and $\bar{u}>0$. The optimal tracking problem is defined as:

@@

\item Second, to force a regularized control. Otherwise note that $u$ is linear in \eqref{Controlsystem}, therefore if the integral cost does not have a non-linear expression of $u$ the optimal control will be of a bang-bang type with possibly singular arcs \cite{harmand2019optimal}. Since the objective is to find a differentiable expression of $I(x)$ the addition of the regularization term is deemed necessary.
\end{itemize} 
The problem of approximating the solution of the system to a desired reference ($z$ in this case) is called the optimal tracking problem. For solving such a problem the approach developed by Cimen \textit{et al.} \cite{Cimen2004, Cimen2008} was adapted to our problem. The method proposed involves the resolution of Approximating Sequences of Ricatti Equations (ASRE). It consists on iteratively calculating trajectories of System \eqref{Controlsystem} with a certain control law and subsequently feed a non-autonomous Ricatti differential equation with the resulting trajectory. Then, a new control law that uses the solution of the Ricatti equation is proposed and a new trajectory is calculated. The iteration is stopped when a convergence in the output or  the control is observed.

The control term should remain positive for the system to be well posed (no negative states), and an upper bound was added to represent the fact that life cannot grow infinitely fast. The tracking problem does not consider a constrained control. Nevertheless, the methods of Cimen \textit{et al.}\cite{Cimen2004} were directly used with an explicit constraint in the synthesis of the control. Even though this is probably suboptimal when the control reaches its bounds, one at least is certain about its optimality when the control never reaches its constraints.

@@

The change of variable $u_i(t) = v_i(t) + 1$ is used for technical reasons explained in the supplementary material section. This in turn implies $v_i(t) \in [-1,\bar{u}-1]$. The feedback control $v(t)$, will be of the form $v(t) = -R^{-1}\tilde{B}^ \top(t)\left(\tilde{P}(t)x(t)-s_f(t)\right)$. Where matrix $P(t)$ and vector $s_f(t)$ solve differential equations. In the appendix it is proved that because of the structure of the system one only needs to calculate $2n$ differential equations for the synthesis of the control, instead of $(n+3)^2 + (n+3)$ that would imply the direct application of the method, which renders the method- at least theoretically- scalable for a growing number of OTU.

@@

The tracking problem was applied to data coming from a nitrification process with experimental conditions described in \cite{dumont2008observers}. For exploring the hypothesis of interactions as drivers of bioreactors performance environmental conditions should be kept as constant as possible. Therefore only data from day $183$ onwards was used because a change in the operating temperature happened at that point, which is known to have an effect on kinetics. For choosing which species belong in which functional group, the procedure described of Ugalde-Salas \textit{et al.} \cite{Ugalde-Salas2019} was used. From day $183$ to day $315$ 31 OTU were identified in the $G_1$ group (AOB) and 5 in the $G_2$ functional group (NOB).

@@

The knowledge of functions $f_i(s)$ was based on a study of nitrification's kinetic parameters\cite{Wiesmann1994}. Particularly given the system's ammonium and nitrate concentration a Monod function (eq \eqref{monod}) was used for $G_1$ and $G_2$ with parameters given in table \ref{kinetic_parameters_application} calculated from the equation of Table 2 of the same article. The yields were fitted to match the nitrogen mass balances. The $Q$ and $R$ matrices were the same as in the proof of concept section, that is $\begin{bmatrix}
\lambda_1 I_{n_1} &0  \\ 0& \lambda_2 I_{n_2}
\end{bmatrix}$ and $I_n$, respectively, with $\lambda_1 = 10^{-4}$ and $\lambda_2 = 10^{-5}$.

@@

\begin{table}[ht]
	\centering
	\begin{tabular}{|l|l|l|l|}
		\hline
		Kinetic Parameters & $\mu_i\,[1/day]$ & $K_i\,[g/L]$ & $\frac{1}{y_i} \, [gr/gr]$ \\ \hline
		$x_1 \in G_1$ & 1.97  & $7\cdot 10^{-1}$ & 4.49  \\ \hline
		$x_2\in G_2$ & 1.87 & $5.4\cdot 10^{-1}$ &  45.51 \\ \hline
	\end{tabular}	
	\caption{A set of kinetic parameters of model \eqref{Controlsystem}.}
	\label{kinetic_parameters_application}
\end{table}
For the reader to gain understanding of the situation, a simulation of the system using the experiments operating parameters ($D$ and $s_{in}$) is presented without control (i.e. $u(t)= 1$) in  figure \ref{u=1 simulation.} nitrite ($s_3$) accumulates all along the trajectory, but when compared to data it is clear that $s_3$ stops accumulating after a while.

@@

Last decades advances in genetic sequencing and microbial ecology have opened a gap for modellers in biochemical processes to integrate this valuable information. Considering the success of mass balance models to predict and pilot bioreactors, one should build upon these existing models. This article proposed a naïve approach to start filling the gap of microbial ecology and biochemical process modelling: The approach can be simply described as correcting the growth rate of each individual species from genetic sequencing measurements, . If one already posses a model for describing a particular bioprocess the use of the techniques here presented can be applied in a straightforward manner if one has already identified the microbial species involved with a particular functionality of the system.

@@

\subsection{Proofs of \cref{l1} and \cref{theoWellPosedness}}

\systembounds*

\begin{proof}
	Define 
	\begin{align}
	z_1 &: = \sum \limits_{i \in G_1} \frac{1}{y_i}x_i + s_1  \\
	z_2 &: = \sum \limits_{i \in G_2} \frac{1}{y_i}x_i + s_1 + s_2 \\
	z_3 &: = s_1 + s_2 + s_3
	\end{align}

@@

\wellposedness*

@@

	If $s_2 = 0$ then note first that by adding both inequalities from Lemma \eqref{l1}:
	\begin{align*}
	\sum \limits_{i \in G_1 \cup G_2}  \frac{1}{y_i} x_i + 2s_1 + s_2 \leq M_1 + M_2 
	\end{align*}
	let $\bar{k} = \min \left\{ \frac{1}{y_i}  : \; i\in [n] \right \}$ and since $s_1,x_i \geq 0$ one has:
	\begin{align*}
	\Rightarrow \bar{k} \sum  \limits_{i = 1}^{n}   x_i  \leq M_1 + M_2  \\
	\Leftrightarrow \bar{k} \sum  \limits_{i = 1}^{n}   \vert x_i \vert  \leq M_1 + M_2  \\
	\Leftrightarrow \bar{k} \Vert x \Vert_1  \leq M_1 + M_2  \\
	\Rightarrow   \bar{k} \Vert x \Vert_{\infty} \leq M_1 + M_2
	\end{align*}

	Define $M := \dfrac{\bar{k}}{M_1 + M_2}$, and let $A$ be a matrix such that $\Vert A \Vert_{\infty} \leq M$. Since
	\begin{align*} \Vert Ax \Vert_{\infty} \leq 1 \Rightarrow   (1 + A_{i\bullet}x) \geq 0 \quad \forall i \in [n]
	\end{align*}

	and $\Vert Ax \Vert_{\infty} \leq \Vert A \Vert_{\infty} \Vert x\Vert_{\infty} \leq M \frac{1}{M}$. One has proven that:
	\begin{align}
	\label{Bound1}  \;   (1 + A_{i\bullet}x) \geq 0  \quad \forall i \in [n]
	\end{align}

@@

	For the boundedness it suffices to notice from \cref{l1} that the sum of positive elements is bounded, therefore each element is bounded.
\end{proof}

\subsection{Deduction of  Equilibrium points}

Recall that

\begin{align}
f_i(s) = \begin{cases}
\bar{\mu}_i \dfrac{s_1}{K_i + s_1} \quad \forall i \in G_1 \\
\bar{\mu}_i \dfrac{s_2}{K_i + s_2} \quad \forall i \in G_2
\end{cases}
\end{align}

then $\mu(x,s) = \diag(f(s))(1_{n\times 1} + Ax)$ thus, system \eqref{system} is rewritten as follows.
